An Automatic Verification Framework for Software Configurations

### 2016-05-02 OSDI/abs.tex
config: for osdi16
config: abs

\begin{abstract}

Software failures resulting from configuration errors 
have become commonplace as modern software systems 
grow increasingly large and more complex.
The lack of language constructs in configuration files, 
such as types and grammars, 
has directed the focus of a configuration file verification 
towards building post-failure error diagnosis tools. 
In addition, the existing tools are generally language
specific, requiring the user to define at least a grammar for the language
models and explicit rules to check. 
In this paper, we propose a framework which analyzes
datasets of correct configuration files and derives rules for building a
language model from the given dataset. 
The resulting language model can be used
to verify new configuration files and detect errors in them. 
Our proposed framework is highly modular, 
does not rely on the system source code, and
can be applied to any new configuration file type with minimal user input.  
Our tool, named \app, relies on an abstract representation of language 
rules to allow for this modularity. 
\app supports learning of various rules, such as orderings,
value relations, type errors, or user defined rules
by using a probabilistic type inference strategy and
defining a small interface for the rule type.

\end{abstract}

>>>>

\section*{Abstract}

System failures resulting from configuration errors 
have become the major factors affecting the availability and
reliability of today's software systems.
While many misconfiguration handling techniques,
\eg, misconfiguration checking and troubleshooting, have been proposed, 
offering automatic verification to configuration files -- like
what we did to regular programs -- is still an open problem.
The main reason is software configuration files are poorly structured 
and typed, thus obstructing any possible language-based verification. 
This paper presents, \app, the first automatic verification framework for
software configurations.
In order to verify a configuration file,
\app first analyzes a given dataset sample of configuration files, 
and then derives rules for building a language model from 
the given configuration dataset. 
Finally, the resulting language model can be used
to verify new configuration files and detect errors in them.
At the heart of \app lies a novel learning engine, 
which employs a probabilistic type inference strategy and
defines an intermediate representation for the rule type.
Using this engine, \app learns various type rules (\eg, orderings,
value relations, type errors, and user defined rules)
to verify configuration files of interest.
\app framework is highly modular, 
does not rely on the system source code, and
can be applied to any new configuration file type with minimal user input.  
%Our tool, named \app, relies on an abstract representation of language 
%rules to allow for this modularity. 
\ennan{One or two sentences here describe the most important
experimental results.}

### 2016-05-02 OSDI/intro.tex
config: for osdi16
config: abs

\section{Introduction and System Overview}
\label{sec:Intro}

@@

\begin{figure}[t] \centering
\includegraphics[width=0.63\textwidth]{figs/overview}
\caption{{\footnotesize \app's workflow. The green boxes represent configuration files,
  including both correct general configuration files and users' input
  configuration files. The purple boxes are the components within \app.
  The yellow boxes are results generated by \app's components.}}
\label{fig-overview}
\end{figure}

Figure~\ref{fig-overview} describes an overview of our system. We start
with the assumption that we are given a number of correct configuration
files belonging to the same category (for instance, MySQL or
Apache). Such files follow similar patterns, which we exploit
in a learning algorithm to build rules that
describe a language model for the files. Since the
``language'' of configuration types is untyped and unstructured, we
first parse the files and translate them into a more structured,
intermediary representation.
When running type inference on a configuration file, the type of a variable cannot always be fully determined from a single value.
We address this problem by introducing so called {\emph{probabalistic types}}.
Rather than giving a variable a single type, we assign several types with their probability distributions. 
We can then use these more structured files
as a training set to learn the rules. The learning algorithm
is template-based to be easily extensible. We provide an initial set of templates and the
learner learns some concrete instances from the training set. These
rules are used for detecting errors violating the learned constraints
in the files given by the user.

>>>>

\section{Introduction}
\label{sec-intro}

@@

Building a practical framework for checking the type-safety of
configuration files, nevertheless, requires addressing several
challenges. First, given the fact that 
different systems' configuration files use various representations, 
the cost of introducing a completely new configuration
language (or representation) to instead of all these existing format looks 
like an impractical goal. This is not only because administrators need 
to learn this new language, but also changes existing system
infrastructure to support such a new language. 
In order to avoid the above issues, we propose a representation,
which is a type-based representation and 
much more structured than existing configuration format, 
but use it as an intermediate layer. We, at the same time,
develop many pluggable parsers
that can transform different systems' configuration representations
into our proposed intermediate representation for the post process.

Second, when running type inference on any given configuration file, 
the type of a variable cannot always be fully determined from a single 
value. For example, an entry {\tt foo = MAX\_SIZE} is most likely
an integer rather than string. We address this problem by introducing 
so called {\emph{probabilistic types}}.
Rather than assigning only one variable to a single type, 
we assign several types with their probability distributions. 
The entry in the above example should be assigned 
a probabilistic type like 
{\tt \{foo, MAX\_SIZE, [(Int, 90\%), (String, 10\%)]\}}.
\ennan{I am not quite sure probabilistic types are obviously
better than deterministic types.}

We can then use these more structured files
as a training set to learn the rules. The learning algorithm
is template-based to be easily extensible. We provide an initial set of templates and the
learner learns some concrete instances from the training set. These
rules are used for detecting errors violating the learned constraints
in the files given by the user.

### 2016-05-02 OSDI/rel.tex
config: abs

Several machine learning-based misconfiguration detection efforts 
also have been proposed~\cite{yuan11context, zhang14encore}.
Specifically, EnCore~\cite{zhang14encore} introduces a template-based
learning approach to improve the accuracy of their learning results.
The learning process is guided by a set of predefined rule templates
that enforce learning to focus on patterns of interests.
By this way, EnCore filters out irrelevant information and reduces
the false positives; moreover, the templates are able to express
system environment information that state-of-the-art machine learning
techniques cannot handle.
Compared with EnCore, \app has the following advantages.
First, \app does not rely on whether the files in given configuration set 
are 100\% correct. Second, \app can cover much more types of 
misconfigurations, since \app introduces probabilistic type checking
mechanism. Finally, \app is a language framework, which could 
even be used to write configuration files, but EnCore is only a 
misconfiguration detection tool.

>>>>

Several machine learning-based misconfiguration detection efforts 
also have been proposed~\cite{yuan11context, zhang14encore}.
EnCore~\cite{zhang14encore} is the most close work to \app.
It introduces a template-based
learning approach to improve the accuracy of their learning results.
The learning process is guided by a set of predefined rule templates
that enforce learning to focus on patterns of interests.
By this way, EnCore filters out irrelevant information and reduces
the false positives; moreover, the templates are able to express
system environment information that other machine learning
techniques cannot handle.
Compared with EnCore, \app has the following advantages.
First, \app does not rely on whether the files in given configuration set 
are 100\% correct. Second, \app not only can cover much more types of 
misconfigurations, but also introduces probabilistic type.
Finally, \app is a language framework, which could 
even be used to write configuration files, but EnCore is only a 
misconfiguration detection tool.
